cardinalR: Generating interesting high-dimensional data structures

### 2025-06-24 paper-cardinalR.Rmd
Updated gen_linearbranches text

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches $1$ and $2$ are initialized with fixed slopes and intercepts. The Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. Branches $3$ to $k$ are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. $X_1$ values span a short segment ($x_{start}$ to $x_{start} + 1$) and $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$ where $s_i$ is a sampled slope from a filtered set and $\epsilon \sim U(0, 0.2)$. This setup yields $k$ spatially distinct, noisy linear trajectories.

For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.

@@

The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in a $p$-dimensional space. The underlying geometry lies in the first two dimensions, while the remaining $p - 2$ dimensions contain Gaussian noise.

>>>>

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches $1$ and $2$ are initialized with fixed slopes and intercepts. The Branch $1$ is generated from $X_1 \sim U(-2, 8)$, $X_2 = 0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. The Branch $2$ is generated from $X_1 \sim U(-6, 2)$, $X_2 = -0.5X_1 + \epsilon$, where $\epsilon \sim U(0, 0.5)$. Branches $3$ to $k$ are added iteratively. Each additional branch $i$ starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. The $X_1$ values are defined over a short range, from $x_{start}$ to $x_{start} + 1$. The $X_2$ value is calculated using the formula $X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon$, where $s_i$ is a chosen slope from a selected branch, and $\epsilon \sim U(0, 0.2)$. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.05^2)$ is added to embed the $2\text{-}D$ branches into $p\text{-}D$, where $j = 3, \dots, p$.

@@

The `gen_curvybranches(n, p, k)` function generates a dataset of $n$ points forming $k$ curvilinear branches embedded in $p\text{-}D$. The underlying geometry lies in the first two dimensions, while the remaining $(p-2)\text{-}D$ dimensions contain Gaussian noise.

### 2025-06-24 paper-cardinalR.tex
Updated gen_linearbranches text

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts. The Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). Branches \(3\) to \(k\) are added iteratively. Each starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. \(X_1\) values span a short segment (\(x_{start}\) to \(x_{start} + 1\)) and \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\) where \(s_i\) is a sampled slope from a filtered set and \(\epsilon \sim U(0, 0.2)\). This setup yields \(k\) spatially distinct, noisy linear trajectories.

For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).

@@

The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in a \(p\)-dimensional space. The underlying geometry lies in the first two dimensions, while the remaining \(p - 2\) dimensions contain Gaussian noise.

>>>>

Each branch is a segment of a line with added jitter to simulate measurement noise. The branches differ in direction and location. Branches \(1\) and \(2\) are initialized with fixed slopes and intercepts. The Branch \(1\) is generated from \(X_1 \sim U(-2, 8)\), \(X_2 = 0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). The Branch \(2\) is generated from \(X_1 \sim U(-6, 2)\), \(X_2 = -0.5X_1 + \epsilon\), where \(\epsilon \sim U(0, 0.5)\). Branches \(3\) to \(k\) are added iteratively. Each additional branch \(i\) starts at a location outside predefined exclusion zones to avoid overlap with the initial two branches. The \(X_1\) values are defined over a short range, from \(x_{start}\) to \(x_{start} + 1\). The \(X_2\) value is calculated using the formula \(X_2 = s_i(X_1 - x_{start}) + y_{start} + \epsilon\), where \(s_i\) is a chosen slope from a selected branch, and \(\epsilon \sim U(0, 0.2)\). For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.05^2)\) is added to embed the \(2\text{-}D\) branches into \(p\text{-}D\), where \(j = 3, \dots, p\).

@@

The \texttt{gen\_curvybranches(n,\ p,\ k)} function generates a dataset of \(n\) points forming \(k\) curvilinear branches embedded in \(p\text{-}D\). The underlying geometry lies in the first two dimensions, while the remaining \((p-2)\text{-}D\) dimensions contain Gaussian noise.
